---
title: "Practical learning machine - Final project"
author: "Manuela Escobar"
Date: October 23 2020
output:
  word_document: default
  html_document: default
---
### Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. In this project, our goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants, and the participants were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

The goal of our project is to predict the manner in which they did the exercise. This is the "class" variable in the training set. We may use any of the other variables to predict with. We should create a report describing how we built our model, how we used cross validation, what we think the expected out of sample error is, and why we made the choices we did. We will also use our prediction model to predict 20 different test cases.

##### Download libraries and reading data
```{r results='hide'}
library(ggplot2)
library(lattice)
library(caret)
library(rpart)
library(RColorBrewer)
library(rattle)
library(e1071)
library(randomForest)
library(mlbench)
library(parallel)
library(doParallel) 
library(corrplot)
library(gbm)
library(dplyr)
train <- read.csv('./data/pml-training.csv', header=T)
validation <- read.csv('./data/pml-testing.csv', header=T)
dim(train)
dim(validation)
```
##### Cleaning data from NA values
```{r}
traindb<- Filter(function(x) !any(is.na(x)), train)
validationdb <- Filter(function(x) !any(is.na(x)), validation)
traindb <- as.data.frame(traindb)
validationdb <- as.data.frame(validationdb)
dim(traindb)
dim(validationdb)
```
##### Making train data base of the same length of validation base
```{r}
classe <- traindb$classe
trainRemove <- grepl("^X|timestamp|window", names(traindb))
traindb <- traindb[, !trainRemove]
trainCleaned <- traindb[, sapply(traindb, is.numeric)]
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(validationdb))
validationdb <- validationdb[, !testRemove]
validCleaned <- validationdb[, sapply(validationdb, is.numeric)]
dim(trainCleaned)
dim(validCleaned)
```
##### Creating test data set
```{r}
traindb <- trainCleaned[createDataPartition(trainCleaned$classe, p=0.7, list=FALSE), ]
testdb <- trainCleaned[-(createDataPartition(trainCleaned$classe, p=0.7, list=FALSE)), ]
nrow(traindb)
nrow(testdb)
```
##### Trainning the model with random forest technique and decission trees
```{r}
cluster <- makeCluster(detectCores() - 1) 
registerDoParallel(cluster)
fitControl <- trainControl(method = "cv", number = 2, allowParallel = TRUE)
system.time(modFit <- train(classe ~ ., method="rf",data=traindb,trControl = fitControl))
stopCluster(cluster)
registerDoSEQ()

predict <- predict(modFit, newdata=testdb)

vi <- varImp(modFit)$importance
vi[head(order(unlist(vi), decreasing = TRUE), 5L), , drop = FALSE]
```
##### Predicting values for validation data
```{r}
finalprediction <- predict(modFit, newdata=validCleaned)
finalprediction
```
##### Figures
```{r}
corrPlot <- cor(traindb[, -length(names(traindb))])
corrplot(corrPlot, method="color")
```
##### Source of the data
Velloso,Bulling, Gellersen,Ugulino, And Fuks. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI, Germany: ACM SIGCHI, 2013.